{\steve}: A hierarchical Bayesian model for Supernova Cosmology

### 2018-12-19 papers/methods/paper.tex
Responding to referee comments

\subsubsection{Uncertainty Propagation}
\label{sec:systreat}
The chief difficulty with including systematic uncertainties in supernova analyses is that they have difficult-to-model effects on the output observations. As such, the traditional treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{{\rm d} \hat{m}_B}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{x}_1}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{c}}{{\rm d} \Z_i}$, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.

@@

This method of creating a secondary covariance matrix using partial derivatives is used throughout the traditional and BHM analyses. For each survey and band, we have two systematics --- the calibration uncertainty and the filter wavelength uncertainty. We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty. A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details. This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey. With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
\begin{equation}
P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C_\eta\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right). \label{eq:l3}
\end{equation}

@@

We thus include into our model parametrised selection effects by including the covariance matrix of selection effect uncertainty. Formally, we include deviations from the determined mean selection function parameters with parameter vector $\Delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.  Whilst this uncertainty encapsulates the potential error from the simulations not matching the analytic approximations, it does not cover potential variations of the selection function at the top level --- varying cosmology or spectroscopic efficiency. Tests with changing the intrinsic scatter model used in the selection efficiency simulations show that the uncertainty introduced is negligible.

With the well-sampled redshift approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova. For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SN~Ia} becomes
\begin{eqnarray}
d_{\rm CDF} &=& \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &=& 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad& \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
\end{eqnarray}
and is incorporated into our likelihood. This is illustrated in Figure \ref{fig:efficiency}. Our corrections for the DES spectroscopic data utilise the CDF functional form, with the combined low redshift surveys being modelled with the skew normal efficiency. Further details on this choice are given in Section \ref{sec:simdes}.

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{plot_pop_simple.pdf}
	\end{center}
	\caption{Population distributions shown in redshift and uncorrected absolute magnitude $m_B - \mu$ for 1000 supernovae in both high-redshift and low-redshift surveys. Selection effects are visible in both samples, where red supernovae are often cut as redshift increases, creating a skewed color population. The color of the data points is representative of the supernova color itself, a negative color value showing bluer supernovae, with positive color values representing redder supernovae.}
	\label{fig:simple_pop}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelOl.pdf}
	\end{center}
	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $\Lambda$CDM model, with a representative contour from a single data realisation shown for context. Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ or $\Omega_\Lambda$ separately due to the severe degeneracy between the parameters.}
	\label{fig:simple_ol}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelW.pdf}
	\end{center}
	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $w$CDM model, with a representative contour from a single data realisation shown for context. The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing misleading statistics due to the non-Gaussian nature of the posterior surface. The recovered posteior maximums show the same degeneracy direction as the representative surface, and scatter around the truth values input into the simulation, which are shown in dashed lines.}
	\label{fig:simple_w}
\end{figure}

@@

\begin{enumerate}[1.]
	\item Draw a redshift from a power law distribution. For the low redshift survey this is $\mathcal{U}(0.0004, 0.01)^{0.5}$, and for the DES-like survey this is $\mathcal{U}(0.008, 1.0)^{0.3}$.
	\item Draw a random mass probability from $\mathcal{U}(0, 1)$ and calculate the mass-brightness correction using $\delta(0) = 0.08$, $\delta(0)/\delta(\infty) = 0.5$, and equation \eqref{eq:mass}.
	\item Draw an absolute magnitude, stretch and color from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$. Calculate the true absolute magnitude using Equation \eqref{eq:standard}.
	\item Calculate $\mu(z)$ given the drawn redshift and cosmological parameters $\Omega_m = 0.3$, $w = -1$ under Flat $\Lambda$CDM cosmology. Use this to determine the true apparent magnitude of the object $m_B$.
	\item Determine if the SN~Ia is detected using detection probability $P(S|m_B) = \mathcal{N}^{\rm skew}(13.72, 1.35, 5.87)$ for the low redshift survey (numeric values obtained by fitting to existing low redshift data). For the DES-like survey, accept with probability $P(S|m_B) = \Phi^C(23.14, 0.5)$. Repeat from step one until we have a supernova that passes. We use realistic values for the selection probability to ensure our model is numerically stable with highly skewed selection functions.
	\item Add independent, Gaussian observational error onto the true $m_B$, $x_1$, $c$ using Gaussian widths of $0.04$, $0.2$, $0.03$ respectively (following the mean uncertainty for DES-like SNANA simulations). Add extra color uncertainty in quadrature of $\kappa_0 + \kappa_1 z$, where $\kappa_0 = \kappa_1 = 0.03$.
\end{enumerate}

@@

To investigate biases in the model in fine detail, we look for systematic bias in $\Omega_m$ in the Flat $\Lambda$CDM cosmology test, and bias in $w$ for the Flat $w$CDM test with strong prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$. This allows us to investigate biases without the investigative hindrances of non-Gaussian or truncated posterior surfaces. The strong prior on $\Omega_m$ cuts a slice through the traditional `banana' posterior surface in the $w$-$\Omega_m$ plane of Figure \ref{fig:simple_w}. Without making such a slice, the variation in $w$ is larger due to a shift along the degeneracy direction of the `banana'. By focusing the slice at an almost fixed $\Omega_m$, we can see the variation in the mean value of $w$ approximately perpendicular to the lines of degeneracy, instead of along them. The results of the analysis are detailed in Table \ref{tab:simple_model}, and demonstrate the performance of our model in recovering the true cosmological parameters.  As we are using 100 independent realisations, the precision of our determination of the mean simulation result is a tenth of the quoted scatter. The deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model. With this simple data, we also correctly recover underlying supernova populations, which can be seen in Figure \ref{fig:simple_w_super}.

@@

Many BHM methods have previously been validated on data constructed explicitly to validate the assumptions of the model. This is a useful consistency check that the model implementation is correct, efficient and free of obvious pathologies. However, the real test of a model is its application to realistic datasets that mimic expected observational data in as many possible ways. To this end, we test using simulations (using the SNANA package) that follow the observational schedule and observing conditions for the DES and low-$z$ surveys, where the low-$z$ sample is based on observations from CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}. Simulation specifics can be found in \citet{Kessler18}.

Additionally, prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach. In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models via simulations, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}. The {\gten} models dispersion with 70\% contribution from coherent variation and 30\% from chromatic variation whilst the {\celeven} model has 25\% coherent scatter and 75\% from chromatic variation. These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.

In addition to the improvements in testing multiple scatter models, we also include peculiar velocities for the low-$z$ sample, and our full treatment of systematics as detailed in \citet{Brout18SYS}. Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}. Initial tests were also done with a second, Gaussian population with color and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency. The selection effects were quantified by comparing all the generated supernovae to those that pass our cuts, as shown in Figure~\ref{fig:sf_snana}. It is from this simulation that our analytic determination of the selection functions for the low-$z$ and DES survey are based. We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the functional form of the Malmquist bias between the two models.

@@

We investigate the cosmological bias and find its source to be a bias in the observed summary statistics (i.e. the $\hat{m}_B$, $\hat{x}_1$, $\hat{c}$ output from SALT2 light curve fitting), in addition to incorrect reported uncertainty on the summary statistics. To confirm this, we run two tests. The first of which, we replace the SALT2-fitted $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a Gaussian centered on the true SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits. With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly. Our second test aims to test our model whilst allowing biases in the summary statistics not caused from intrinsic scatter through. To this end, we test a set of 100 simulations generated using an intrinsic dispersion model of only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model. From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased. Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced by an unknown scatter model.

@@

Difficulty in providing an adequate parametrisation for realistic intrinsic dispersion, and the simplification of Malmquist bias to only apparent magnitude also leads to biases in the population parameters. Tests when fixing the population parameters to true does not resolve the cosmological bias observed in the {\celeven} simulations. As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.

@@

With the increased flexibility of Bayesian hierarchical models over traditional models, we expect to find an increased uncertainty on parameter inference. To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES-SN3YR sample with various model parameters allowed to either vary or stay locked to a fixed value. By taking the difference in uncertainty in quadrature, we can infer the relative contribution for each model feature to the uncertainty error budget.

>>>>

\subsubsection{Uncertainty Propagation}
\label{sec:systreat}
The chief difficulty with including systematic uncertainties in supernova analyses is that they have difficult-to-model effects on the output observations. As such, the traditional treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{{\rm d} \hat{m}_B}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{x}_1}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{c}}{{\rm d} \Z_i}$ {\blue for each supernova light curve fit}, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.

@@

This method of creating a secondary covariance matrix using partial derivatives is used throughout the traditional and BHM analyses. For each survey and band, we have two systematics --- the calibration uncertainty and the filter wavelength uncertainty. We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty. A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details. This gives thirteen global systematics shared by {\blue all surveys that apply globally to all supernova summary statistics}, plus two systematics per band in each survey. With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
\begin{equation}
P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C_\eta\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right). \label{eq:l3}
\end{equation}

@@

We thus include into our model parametrised selection effects by including the covariance matrix of selection effect uncertainty. Formally, we include deviations from the determined mean selection function parameters with parameter vector $\delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.  Whilst this uncertainty encapsulates the potential error from the simulations not matching the analytic approximations, it does not cover potential variations of the selection function at the top level --- varying cosmology or spectroscopic efficiency. Tests with changing the intrinsic scatter model used in the selection efficiency simulations show that the uncertainty introduced is negligible.

With the well-sampled redshift approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova. For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SN~Ia} becomes
\begin{eqnarray}
d_{\rm CDF} &=& \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &=& 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad& \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
\end{eqnarray}
and is incorporated into our likelihood. {\blue Note that the above efficiencies utilise the common form of the normal distribution rather than the conditional probability notation found previously in this work.} This is illustrated in Figure \ref{fig:efficiency}. Our corrections for the DES spectroscopic data utilise the CDF functional form, with the combined low redshift surveys being modelled with the skew normal efficiency. Further details on this choice are given in Section \ref{sec:simdes}.

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{plot_pop_simple.pdf}
	\end{center}
	\caption{Population distributions shown in redshift and uncorrected absolute magnitude $m_B - \mu$ for 1000 supernovae in both high-redshift and low-redshift surveys. Selection effects are visible in both samples, where red supernovae are often cut as redshift increases, creating a skewed color population. The color of the data points is representative of the supernova color itself, a negative color value showing bluer supernovae, with positive color values representing redder supernovae.}
	\label{fig:simple_pop}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelOl.pdf}
	\end{center}
	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $\Lambda$CDM model, with a representative contour from a single data realisation shown for context. Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ or $\Omega_\Lambda$ separately due to the severe degeneracy between the parameters.}
	\label{fig:simple_ol}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelW.pdf}
	\end{center}
	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $w$CDM model, with a representative contour from a single data realisation shown for context. The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ {\blue shown to recover input cosmology. For contours that are non-Gaussian due to the curved degeneracy between $\Omega_m$ and $w$, the marginalised distributions can provide misleading statistics where maximum marginalised distribution can disagree with the maximum likelihood in multiple dimensions. For our contours, the non-Gaussianity is small and the marginalised distributions still provide a valuable metric.} The recovered posteior maximums show the same degeneracy direction as the representative surface, and scatter around the truth values input into the simulation, which are shown in dashed lines.}
	\label{fig:simple_w}
\end{figure}

@@

\begin{enumerate}[1.]
	\item Draw a redshift from a power law distribution. For the low redshift survey this is $\mathcal{U}(0.0004, 0.01)^{0.5}$, and for the DES-like survey this is $\mathcal{U}(0.008, 1.0)^{0.33}$. {\blue For the low redshift survey, this is equivalent to sampling $y = z$ from $0.02$ to $\sqrt{0.1}$, for the high redshift survey, this is equivalent to sampling $y = z^2$ from $0.2$ to $1.0$. These distributions are arbitrary, and this test has been performed with various flat and power law distributions.}
	\item Draw a random mass probability from $\mathcal{U}(0, 1)$ and calculate the mass-brightness correction using $\delta(0) = 0.08$, $\delta(0)/\delta(\infty) = 0.5$, and equation \eqref{eq:mass}.
	\item Draw an absolute magnitude, stretch and color from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$.
	\item Calculate $\mu(z)$ given the drawn redshift and cosmological parameters $\Omega_m = 0.3$, $w = -1$ under Flat $\Lambda$CDM cosmology. Use this to determine the true apparent magnitude of the object $m_B$ using $m_B = \mu + M_B - \alpha x_1 + \beta c$.
	\item Determine if the SN~Ia is detected using detection probability $P(S|m_B) = \mathcal{N}^{\rm skew}(13.72, 1.35, 5.87)$ for the low redshift survey (numeric values obtained by fitting to existing low redshift data). For the DES-like survey, accept with probability $P(S|m_B) = \Phi^C(23.14, 0.5)$. Repeat from step one until we have a supernova that passes. We use realistic values for the selection probability to ensure our model is numerically stable with highly skewed selection functions.
	\item Add independent, Gaussian observational error onto the true $m_B$, $x_1$, $c$ using Gaussian widths of $0.04$, $0.2$, $0.03$ respectively (following the mean uncertainty for DES-like SNANA simulations). Add extra color uncertainty in quadrature of $\kappa_0 + \kappa_1 z$, where $\kappa_0 = \kappa_1 = 0.03$.
\end{enumerate}

@@

To investigate biases in the model in fine detail, we look for systematic bias in $\Omega_m$ in the Flat $\Lambda$CDM cosmology test, and bias in $w$ for the Flat $w$CDM test with strong prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$. This allows us to investigate biases without the investigative hindrances of non-Gaussian or truncated posterior surfaces. The strong prior on $\Omega_m$ cuts a slice through the traditional `banana' posterior surface in the $w$-$\Omega_m$ plane of Figure \ref{fig:simple_w}. Without making such a slice, the variation in $w$ is larger due to a shift along the degeneracy direction of the `banana'. By focusing the slice at an almost fixed $\Omega_m$, we can see the variation in the mean value of $w$ approximately perpendicular to the lines of degeneracy, instead of along them. The results of the analysis are detailed in Table \ref{tab:simple_model}, and demonstrate the performance of our model in recovering the true cosmological parameters.  As we are using 100 independent realisations, the precision of our determination of the mean simulation result is {\blue approximately a tenth of the quoted scatter (as a degree of non-Gaussianity of our fits will make this relationship inexact)}. The deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model. With this simple data, we also correctly recover underlying supernova populations, which can be seen in Figure \ref{fig:simple_w_super}.

@@

Many BHM methods have previously been validated on data constructed explicitly to validate the assumptions of the model. This is a useful consistency check that the model implementation is correct, efficient and free of obvious pathologies. However, the real test of a model is its application to realistic datasets that mimic expected observational data in as many possible ways. To this end, we test using simulations (using the SNANA package) that follow the observational schedule and observing conditions for the DES and low-$z$ surveys, where the low-$z$ sample is based on observations from CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}. Simulation specifics can be found in \citet{Kessler18}. {\blue The primary difference between the toy data of the previous section are the different underlying colour and stretch, inclusion of spectroscopic data cuts and light curve cuts, and inclusion of intrinsic dispersion models.}

{\blue The inclusion of a proper treatment of intrinsic scatter is significant, as} prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach. In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models via simulations, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}. The {\gten} models dispersion with 70\% contribution from coherent variation and 30\% from chromatic variation whilst the {\celeven} model has 25\% coherent scatter and 75\% from chromatic variation. These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.

In addition to the improvements in testing multiple scatter models, we also include peculiar velocities for the low-$z$ sample, and our full treatment of systematics as detailed in \citet{Brout18SYS}. Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}. Initial tests were also done with a second, Gaussian population with color and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency. The selection effects were quantified by comparing all the generated supernovae to those that pass our cuts, as shown in Figure~\ref{fig:sf_snana}. It is from this simulation that our analytic determination of the selection functions for the low-$z$ and DES survey are based. We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the functional form of the Malmquist bias between the two models. {\blue Uncertainty on the analytic selection function is incorporated into our fits, mitigating the imperfection of our analytic form by allowing it to vary in our fits.}

@@

We investigate the cosmological bias and find its source to be a bias in the observed summary statistics (i.e. the $\hat{m}_B$, $\hat{x}_1$, $\hat{c}$ output from SALT2 light curve fitting), in addition to incorrect reported uncertainty on the summary statistics. To confirm this, we run two tests. The first of which, we replace the SALT2-fitted $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a Gaussian centered on the true SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits. With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly. Our second test aims to test our model whilst allowing biases in the summary statistics not caused from intrinsic scatter through. {\blue That is, the first test ascertained that biases in the summary statistics are the cause of cosmological bias. It is thus important to determine the source of those biases; whether they are from intrinsic scatter model or another aspect of the simulation. To} this end, we test a set of 100 simulations generated using an intrinsic dispersion model of only coherent magnitude scatter. {\blue We find $w=-1.00$, showing} that the source of the biases in summary statistics is the underlying intrinsic scatter model. From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased. Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced by an unknown scatter model.

@@

Difficulty in providing an adequate parametrisation for realistic intrinsic dispersion, and the simplification of Malmquist bias to only apparent magnitude also leads to biases in the population parameters. {\blue To determine if the underlying population mischaracterisation was a cause of cosmological bias, we ran fits wherein the underlying population was fixed to the known distributions used for the simulation. These fits did not show reduced cosmological bias in the {\celeven} simulations, and thus the determination of underlying population in our model may give biased distributions, however does not impact cosmology.} As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.

@@

With the increased flexibility of Bayesian hierarchical models over traditional models, we expect to find an increased uncertainty on parameter inference. {\blue This increased uncertainty is one of the strengths of hierarchical models as it represents a more thorough accounting of model uncertainty.} To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES-SN3YR sample with various model parameters allowed to either vary or stay locked to a fixed value. By taking the difference in uncertainty in quadrature, we can infer the relative contribution for each model feature to the uncertainty error budget.
